---
title: "Lab Notebook"
output: html_notebook
---
#Intro
This project is a collaboration between myself and Dr. Roger Day as we observe the Weakest Link Model and its relation to biometrics.  So far, we have plotted with binary and continuous data and plotted using the likelihood function.  This function was used to plot contours that were used to find the most likely values in a given data set.  For now the data in the plots is synthetic, but we will soon use real-world genetic data in plots. 


###WLbinarydata
This code is a function that generates random binary data with 200 "patients", checks the minimum value of each patient(either 0 or 1), and plots each "patient" in pairs based on that value.
```{r}
WLbinarydata <-
function(n = 200, p0 = .1, p1 = .5){
  x1 = rbinom(n = n, size = 1, prob = 1/2)
  x2 = rbinom(n = n, size = 1, prob = 1/2)
  x1x2 = cbind(x1, x2)
  both = apply(X = x1x2, MARGIN = 1, FUN = min)
  yprob = p0 * (both == 0) + p1 * (both == 1)
  y = rbinom(n = n, size = 1, prob = yprob)
  return(data.frame(x1, x2, y, both))
}
```

###WLContinuousdata
This code is a function that generates random continuous data, rather than binary data with 200 "patients".  Using the phi1 and phi2 equations, along with inverse logit, the minimum phi is calculated for each "patient" and plotted in pairs.
```{r}
WLContinuousdata <-
function(n = 200, p0 = .1, p1 = .5, a1 = 1, a2 = 2, b1 = 1, b2 = 2){
  x1 = rnorm(n = n)
  x2 = rnorm(n = n)
  x1x2 = cbind(x1, x2)
  phi1 = a1 + b1 * x1
  phi2 = a2 + b2 * x2
  phi1phi2 = cbind(phi1, phi2)
  minphi = apply(X = phi1phi2, MARGIN = 1, FUN = min)
  yprob = inverselogit(minphi)
  y = rbinom(n = n, size = 1, prob = yprob)
  return(data.frame(x1, x2, y, minphi))
}
```

###Likelydata
This code is a function that simulates the likelihood function using the phi and inverse logit equations from eariler in the program.The phis are generated based on a1,a2,b1, and b2 from the argument and the minimum is put into the likelihood equation.

```{r}
Likelydata <-
function(data = temp, a1 = 1, a2 = 2, b1 = 1, b2 = 2, a1a2b1b2){
  if(!missing(a1a2b1b2)){
    a1 = a1a2b1b2$a1
    a2 = a1a2b1b2$a2
    b1 = a1a2b1b2$b1
    b2 = a1a2b1b2$b2
  }
  phij1 = inverselogit(a1 + b1 * data$x1)
  phij2 = inverselogit(a2 + b2 * data$x2)
  phij1phij2 = cbind(phij1, phij2)
  minphij = apply(X = phij1phij2, MARGIN = 1, FUN = min)
  return(sum(data$y*log(minphij) + (1 - data$y) * log(1 - minphij)))
}
```

####Plotting for Likelydata
The a1 and a2 seqs are then generated, and the a1 seq is plotted at the x axis against the Likely values for one dimension in a curve of optimal use.
```{r}
Likelydata()
a1seq = seq(0,2,length = 100)
a2seq = seq(0,4,length = 100)

Lvalues = sapply(a1seq, Likelydata, data = temp)
plot(a1seq, Lvalues, type = "l")
abline(v = a1seq[which.max(Lvalues)])
abline(h = max(Lvalues))
```

###Two_Dimensional
This code is a function in which a multi-dimensional contour is generated based on the a1seq, a2seq, and Lvalues.  The Lvalues are then taken with the continuous data function and the maximum likelihood is found  and plotted in the middle of the contour based on the Lvalues row and column maximizer.
```{r}
Two_Dimensional = function(){
  a1a2grid = expand.grid(a1seq, a2seq)
  Lvalues = apply(a1a2grid, 1, function(row){Likelydata(a1 = row[1], a2 = row[2])})
  Lvalues = matrix(Lvalues, nrow = 100)
  contour.default(x = a1seq, y =a2seq, z = Lvalues)
  args(WLContinuousdata)
  points(1, 2, col = "green", pch = "x", cex = 2)
  maximizer = which(max(Lvalues) == Lvalues)
  rmax = row(Lvalues)[maximizer]
  cmax = col(Lvalues)[maximizer]
  points(a1seq[rmax], a2seq[cmax], col = "red", pch = "o", cex = 4)
  return(c(maximum = max(Lvalues), a1maximizer = a1seq[rmax], a2maximizer = a2seq[cmax]))
}
```
###More Accurate Maximum Likelihoods
Each dimension is given a function that changes its value from the original starting vector after it moves.  A green circle starts off in a random vector and moves along each dimension in order to find a more accurate maximum likelihood than the on previously found.  The Result is then plotted in  pairs and returned to the user.

```{r}
A1_Varies = function(a1){
  Likelydata(a1 = a1, 
             b1 = starting_vector$b1, 
             a2 = starting_vector$a2, 
             b2 = starting_vector$b2)
}
A2_Varies = function(a2){
  Likelydata(a1 = starting_vector$a1, 
             b1 = starting_vector$b1, 
             a2 = a2, 
             b2 = starting_vector$b2)
}
B1_Varies = function(b1){
  Likelydata(b1 = b1, 
             a1 = starting_vector$a1, 
             a2 = starting_vector$a2, 
             b2 = starting_vector$b2)
}
B2_Varies = function(b2){
  Likelydata(a1 = starting_vector$a1, 
             b1 = starting_vector$b1, 
             b2 = b2, 
             a2 = starting_vector$a2)
}
ConMaxStep = function(){
  newa1 = optimize(f = A1_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$a1 <<- newa1
  newa2 = optimize(f = A2_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$a2 <<- newa2
  newb1 = optimize(f = B1_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$b1 <<- newb1
  newb2 = optimize(f = B2_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$b2 <<- newb2
  return(c(maximum = Likelydata(a1a2b1b2 = starting_vector), 
           a1maximizer = newa1, 
           a2maximizer = newa2,
           b1maximizer = newb1, 
           b2maximizer = newb2))
}
ConMax = function(tol = 1e-7, 
                  starting_vector = list(a1 = 1, a2 = 3, b1 = 0, b2 = 1)
){
  starting_vector <<- starting_vector
  shouldstop = FALSE
  Results = NULL
  while(shouldstop == FALSE){
    Temp = ConMaxStep()
    if(is.null(Results)){
      Results = t(as.data.frame(Temp))
    }
    else{
      Results = rbind(Results, Temp)
      Delta = Results[nrow(Results), "maximum"] -
        Results[nrow(Results) - 1, "maximum"]
      if(Delta < tol){
        shouldstop = TRUE
      }
    }
  }
  return(Results)
}
ConMaxResult = ConMax(tol = 1e-9)
pairs(ConMaxResult)
```

#### Plotting the Line of Efficient Use
This first piece of code is used to generate random points on a plot using the Continuous data, and x1 and x2 variables.  The slope of the plot will then be plotted based on the location of the plot and its points, and the lines of efficeint use (The "L" shapes) will be plotted based on the slope.
```{r}
#The purpose of this function is to plot points on a plane of x1 by x2.
dataset = WLContinuousdata(n = 100, p1 = .9, b1 = 5, b2 = 5)
Plotpoints = function(data = dataset){
  plot(data$x1, data$x2, pch = as.character(data$y), col = c("red", "blue")[1 + data$y])
}
Plotpoints()
attach(as.list(attributes(dataset)[["parameters"]]))
abline(a = ((a1 + a2)/ b2), b = (b1/b2))
detach()
A1_Varies = function(a1){
  Likelydata(a1 = a1, 
             b1 = starting_vector$b1, 
             a2 = starting_vector$a2, 
             b2 = starting_vector$b2)
}
A2_Varies = function(a2){
  Likelydata(a1 = starting_vector$a1, 
             b1 = starting_vector$b1, 
             a2 = a2, 
             b2 = starting_vector$b2)
}
B1_Varies = function(b1){
  Likelydata(b1 = b1, 
             a1 = starting_vector$a1, 
             a2 = starting_vector$a2, 
             b2 = starting_vector$b2)
}
B2_Varies = function(b2){
  Likelydata(a1 = starting_vector$a1, 
             b1 = starting_vector$b1, 
             b2 = b2, 
             a2 = starting_vector$a2)
}
ConMaxStep = function(){
  newa1 = optimize(f = A1_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$a1 <<- newa1
  newa2 = optimize(f = A2_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$a2 <<- newa2
  newb1 = optimize(f = B1_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$b1 <<- newb1
  newb2 = optimize(f = B2_Varies, interval = c(0,2), maximum = TRUE)$maximum
  starting_vector$b2 <<- newb2
  return(list(maximum = Likelydata(a1a2b1b2 = starting_vector), 
           a1maximizer = newa1, 
           a2maximizer = newa2,
           b1maximizer = newb1, 
           b2maximizer = newb2))
}

estimateline = ConMaxStep()
abline(a = (estimateline$a1maximizer - estimateline$a2maximizer)/estimateline$b2maximizer,
       b = estimateline$b1maximizer/estimateline$b2maximizer, 
       col = "purple", 
       lty = 2, 
       lwd = 2) 

```

