---
title: "Pairing PAM50 genes"
runtime: shiny
output: 
  html_notebook: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(WeakestLink)
require(survival)
library(GenSA)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Part A: Choosing/Plotting a dataset

Choose your dataset between METABRIC and Simulated data.  If stimualted data is chosen, adjust the population, as well as the a's and b's.  If METABRIC data is chosen, choose to pair up 45 of the PAM50 genes.
```{r eruptions, echo=FALSE}
rvalues = reactiveValues()
inputPanel(
  radioButtons("chosen_data", label = "Dataset:",
              choices = c("Metabric", "Simulated"), selected = "Simulated"),
  conditionalPanel("input.chosen_data == 'Simulated'",
                   numericInput("sim_population", label = "Simulated Population",
                                min = 0, max = 500, value = 100),
                   sliderInput("a1_adjust", label = "A1 adjustment:",
                               min = 1, max = 10, value = 1, step = 1), 
                   sliderInput("a2_adjust", label = "A2 adjustment:",
                               min = 1, max = 10, value = 2, step = 1), 
                   sliderInput("b1_adjust", label = "B1 adjustment:",
                               min = 1, max = 10, value = 1, step = 1), 
                   sliderInput("b2_adjust", label = "B2 adjustment:",
                               min = 1, max = 10, value = 2, step = 1)
  ), 
  conditionalPanel("input.chosen_data == 'Metabric'",
                   selectInput("Gene1", label = "G1",
                               choices = p45),
                   selectInput("Gene2", label = "G2",
                               choices = p45),
                   selectInput("Endpoints", label = "Endpoints",
                               choices = c("D7")
                   )
                   
  ) 
)

renderPlot({
  if (input$chosen_data == 'Metabric'){
    g1= input$Gene1
    g2= input$Gene2
    print(g2)
    resultFitted = analyzeAPair(g1 = g1, g2 = g2,
                                endpoint = mb$D7,  # Surv(mb$time, mb$cens) 
                                delta = 'fit',
                                plottheData= TRUE)
    drawCOU(x1 = mb[[g1]], x2 = mb[[g2]], delta = 0, FhatStyle = 'normal',
            col='blue')
    drawCOU(x1 = mb[[g1]], x2 = mb[[g2]], delta = 0, FhatStyle = 'ecdf',
            col='purple')
  }
    else{
      rvalues$sim_data = WLContinuousdata(n = input$sim_population, 
                       a1 = input$a1_adjust, 
                       a2 = input$a2_adjust,
                       b1 = input$b1_adjust,
                       b2 = input$b2_adjust)
      fitQWLprobit(rvalues$sim_data, delta = c(-3, 3))
      drawCOUline(a1 = input$a1_adjust, 
                  a2 = input$a2_adjust,
                  b1 = input$b1_adjust,
                  b2 = input$b2_adjust, 
                  col = "purple",
                  lty = 2,
                  lwd = 2)
      
      
    }
  
}
)
```

## Estimation Methods

```{r tabsets, echo=FALSE}
inputPanel(
  checkboxGroupInput("estimethods", label = "Estimation Methods",
                     choices = c("Generated Stimulated Annealing", 
                     "Quantitative Stitching (Delta Fixed)",
                     "Quantitative Stitching (Delta Fitted)",
                     "Conditional Optimization")), 
  uiOutput("estimationplots")
  
)
output$estimationplots = renderUI(expr = {
  if ("Generated Stimulated Annealing" %in% input$estimethods){
    starting_vector = c(a1=0, a2=0, b1=1, b2=1)
    sv = starting_vector
    if (input$chosen_data == "Metabric"){
      mydata = data.frame(x1 = mb[[input$Gene1]], 
                          x2 = mb[[input$Gene2]], 
                          y = mb[[input$Endpoints]])  
    }
    else{
      mydata = rvalues$sim_data
    }
    print(dim(mydata))
    print(names(mydata))
    myfn = function(a1a2b1b2){
      -1 * Likelydata(a1a2b1b2 = a1a2b1b2, data = mydata)
    }
    
    GenSA::GenSA(par = sv, fn = myfn, lower = sv - 3, upper = sv + 3, 
    control = list(maxit = 100))
    drawCOUline(a1 = SA$par["a1"], 
             a2 = SA$par["a2"], 
             b1 = SA$par['b1'], 
             b2 = SA$par["b2"])

  }
if ("Quantitative Stitching (Delta Fixed)" %in% input$estimethods){
  #single number delta numeric input
  #FitQWLprobit 
}
if ("Quantitative Stitching (Delta Fitted)" %in% input$estimethods){
  #Leave delta blank
}
if ("ConditionalMaximum Optimization" %in% input$estimethods){
  #Beginning program
  
}

})
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



